---
title: "p8105_hw2_xw2676.Rmd"
output: github_document
---

## Problem1
#### Read and clean the Mr. Trash Wheel sheet:
```{r setup, include=FALSE}
library(readxl)
library(tidyverse)
library(ggridges)
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
Mr.Trash_Wheel = read_excel("./dataset/HealthyHarborWaterWheelTotals2018-7-28.xlsx",
                            sheet = "Mr. Trash Wheel",
                            range = "A2:N338") %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>%
  mutate(sports_balls = as.integer(round(sports_balls,digits = 0)))
```

#### Read and clean precipitation data for 2017 and 2018:
```{r}
data_2017 = read_excel("./dataset/HealthyHarborWaterWheelTotals2018-7-28.xlsx",
                       sheet = "2017 Precipitation",range = "A2:B14") %>%
  janitor::clean_names() %>% 
  drop_na(total) %>% 
  mutate(year = 2017)
data_2018 = read_excel("./dataset/HealthyHarborWaterWheelTotals2018-7-28.xlsx",
                       sheet = "2018 Precipitation",range = "A2:B14") %>%
  janitor::clean_names() %>% 
  drop_na(total) %>% 
  mutate(year = 2018)
data_2017_2018 = bind_rows(data_2017, data_2018) %>%
  mutate(month=month.name[month])
```
#### Interpretation
There are `r nrow(Mr.Trash_Wheel)` observations,  `r ncol(Mr.Trash_Wheel)` variables in Mr.Trash_Wheel dataset, whose key variables are `r colnames(Mr.Trash_Wheel[5:14])`. 
There are `r nrow(data_2017_2018)` observations,  `r ncol(data_2017_2018)` variables in data_2018 dataset, whose key variables are `r colnames(data_2017_2018[2])`. 

The total precipitation in data_2018 was `r sum(pull(data_2018,total))`.
The median number of sports balls in a dumpster in 2017 was `r median(pull(filter(Mr.Trash_Wheel,year == "2017"),sports_balls))`.

Mr. Trash Wheel dataset includes information on the dumpter number, date of collection, amount of total litter and litter type. 

data_2017_2018 dataset includes monthly precipitation amounts in 2017 and 2018.

## Problem2
#### clean the data in pols-month.csv
```{r}
pols_month = read_csv("./dataset/pols-month.csv") %>%
  janitor::clean_names() %>% 
  separate(mon, into = c("year","month","day"), sep = "-")%>%
  mutate(
    year=as.integer(year),
    month=as.integer(month),
    month=month.name[month])%>% 
  pivot_longer(
    c("prez_gop","prez_dem"),
    names_to = "president",
    names_prefix = "prez_") %>% 
  filter(value == 1|value == 2) %>% 
  select(-value,-day) 
```
#### clean the data in snp.csv
```{r}
snp = read_csv("./dataset/snp.csv") %>%
  janitor::clean_names() %>% 
  separate(date, into = c("month","day","year"), sep = "/") %>%
  mutate(
    year = as.numeric(year),
    month = as.numeric(month),
    month = month.name[month]) %>% 
    select(-day) %>% 
    arrange(year,month)
snp = snp[,c(2,1,3)]
```
#### tidy the unemployment data
```{r}
unemployment = read_csv("./dataset/unemployment.csv") %>%
  janitor::clean_names() %>% 
  pivot_longer(
    jan:dec,
    names_to = "month",
    values_to = "unemployment") %>% 
  mutate(
    month = factor(month, 
                   levels =
  c("jan","feb","mar","apr","may","jun","jul","aug","sep","oct","dec","nov"),
                   labels =
  c("1","2","3","4","5","6","7","8","9","10","11","12"))
  ) %>% 
  mutate(
    month = month.name[as.numeric(month)]
  )
```
#### Join the datasets by merging snp into pols, and merging unemployment into the result.
```{r}
pols_month = unite(pols_month,date,year:month)
snp = unite(snp,date,year:month)
unemployment = unite(unemployment,date,year:month)
join_data = 
  left_join(pols_month, snp, by = "date") %>% 
  left_join(unemployment, by = "date")
```

#### Interpretation
There are `r nrow(pols_month)` observations and `r ncol(pols_month)` variables in pols_month dataset. This dataset is related to the number of national politicians who are democratic or republican at any given time. The key variables are `r colnames(pols_month[2:8])`. 

There are `r nrow(snp)` observations and `r ncol(snp)` variables in snp dataset. The key variables are `r colnames(snp[2])`. This dataset is related to Standard & Poor’s stock market index (S&P), often used as a representative measure of stock market as a whole

There are `r nrow(unemployment)` observations and `r ncol(unemployment)` variables in unemployment dataset. This dataset is the monthly data of  unemployment percentage between 1948 and 1959.The key variables are `r colnames(unemployment[2])`. 

There are `r nrow(join_data)` observations and `r ncol(join_data)` variables in join_data. The key variables are `r colnames(join_data[2:10])`. The range of year is from `r min(pull(join_data, date))` to `r max(pull(join_data, date))`. 

## Problem3
#### Load and tidy the data. 
```{r}
Popular_Baby_Names = read_csv("./dataset/Popular_Baby_Names.csv") %>%
  janitor::clean_names() %>% 
  mutate(childs_first_name = str_to_title(childs_first_name)) %>% 
  distinct() %>% 
  mutate(
    ethnicity = recode(ethnicity, "ASIAN AND PACI" = "ASIAN AND PACIFIC ISLANDER",
                       "BLACK NON HISP" = "BALCK NON HISPANIC",
                       "WHITE NON HISP" = "WHITE NON HISPANIC")
  )
```

#### Produce a well-structured, reader-friendly table showing the rank in popularity of the name “Olivia” as a female baby name over time

```{r}
olivia = filter(Popular_Baby_Names, childs_first_name == "Olivia") %>%
  select(-gender,-childs_first_name,-count) %>% 
pivot_wider(names_from = year_of_birth,
            values_from = rank) %>% 
  knitr::kable(format = "html")
olivia
```

#### Produce a similar table showing the most popular name among male children over time.

```{r}
popular_male_name = filter(Popular_Baby_Names, gender == "MALE",rank == 1) %>% 
  select(-rank, -count, -gender) %>% 
  pivot_wider(
  names_from = "year_of_birth",
  values_from = c("childs_first_name"))
popular_male_name
```

#### produce a scatter plot 
```{r}
male_white_nonhispanic_2016 = filter(Popular_Baby_Names, gender == "MALE", ethnicity == "WHITE NON HISPANIC", year_of_birth == 2016) %>% 
ggplot( aes(x = rank, y = count)) + 
  geom_point()
ggsave("sactterplot of male_white_nonhispanic_2016.pdf")
```















